\prefacesection{Acknowledgements}
I am deeply grateful to my advisor, Steve Quake, for his unwavering belief in his students and his kind guidance. Steve's contributions to the work presented here have been invaluable, as has been his role in my growth as a scientist. The fun, hard-working, and rewarding environment he has cultivated in his lab is something I greatly appreciate. What I admire most about Steve is his broad range of interests and his incisive lines of questioning.

My sincere thanks go to my committee members, Dan Jarosz, Daria Mochly-Rosen, and Bali Pulendran, for their insightful guidance and probing questions. I would also like to express my heartfelt gratitude to Felix Horns. His contribution to my development as a graduate student cannot be overstated. From him, I have learned about single-cell sequencing, the molecular biology underpinning these technologies, and bioinformatics. He taught me how to think critically about data and read scientific literature carefully. His patience as I navigated my learning journey, and his qualities as a collaborator, mentor, and friend, are deeply appreciated.

I extend my thanks to Bob Jones and Sheela Crasta for their help in coordinating sample collections for the Tabula Bursa work, and to the anonymous organ donors who made this work possible. I am grateful to Norma Neff, Rose Yan, and Angela Detweiler for their assistance in generating most of the sequencing data presented here, and to Ivana Cvijovic for being a fantastic collaborator and friend.

I express my gratitude to all members of the Quake group. Despite the high turnover in academic labs, the soul of our lab — always curious and courageous in the pursuit of science — has remained constant. Working with this team has been a joy. I would like to give special thanks to Mira Moufarrej, with whom I shared a bay. Her company made coming to work enjoyable, and her help with computational and experimental issues was invaluable.

I thank Whitney Combes, Vickie Lin, and Sheela Crasta for their roles in keeping the lab running smoothly. My work was supported by an NIH training grant and a National Science Foundation Graduate Research Fellowship, for which I am grateful. Lastly, but by no means least, I thank my friends and family for their support throughout my graduate school journey. In particular, I want to acknowledge my Mom and Dad, the most supportive people I've ever known. I love them dearly.